Off the beaten path: \\ Gaia reveals GD-1 stars perturbed off of the main stream track

### 2018-04-29 paper/gd1.tex
add reminder
Density variations

@@

To study of global properties of the stream as a function of stream longitude,
we extract stream stars around a low-order polynomial model for the median
stream latitude, $\phi_2$, as a function of stream longitude, $\phi_1$.
In detail, we find stream ridge points by computing a running median in $\phi_2$
in steps of $XX^\circ$ along $\phi_1$, then fit a second-order polynomial to the
ridge points to find $\phi_{2, \textrm{track}}(\phi_1)$.
We define the stream as the region within $\left| \phi_2 - \phi_{2,
\textrm{track}}(\phi_1) \right| < 0.6^\circ$.
\figurename~\ref{fig:XX} (XX panel) again shows the high-confidence
stream stars, and the two curved lines (blue) show the upper and lower boundary
of the stream region.

\begin{figure}[h]
\begin{center}
% \includegraphics[width=\textwidth]{???.pdf}
\end{center}
\caption{%
\todo{???}
\label{fig:XX}
}
\end{figure}

The clearly identifiable portions of the GD-1 stream are located at high
Galactic latitudes ($b > 20^\circ$), and we therefore do not expect significant
dust extinction or variations in extinction within the stream region.
\figurename~\ref{fig:XX} (XX panel) shows the $V$-band extinction
in the region around the GD-1 stream, computed from the
Schlegel-Finkbeiner-Davis extinction map (\cite{Schlegel:1998}; hereafter SFD).
For stream longitudes $-60^\circ < \phi_1 < 10^\circ$ and $-1^\circ < \phi_2 <
1^\circ$, the median $V$-band extinction is $\textrm{med}\left(A_V\right) =
0.04~\textrm{mag}$, and the dispersion in $A_V$ in this region is $\sigma_{A_V}
\approx 0.03~\textrm{mag}$.
For longitudes $\phi_1 < -60^\circ$, as the stream approaches the Galactic disk,
dust extinction becomes more appreciable, and the stream stars are harder to
select apart from the background.

We extract properties of the stream from the region defined above as a function
of $\phi_1$ by computing properties in overlapping windows:
we set the window size to $XX^\circ$ in $\phi_1$, and shift the window center by
$1^\circ$ along the stream region.
\figurename~\ref{fig:track-and-model} shows the median latitude, $\phi_2$,
stream surface density, $XX$, stream width, $XX$, and median proper motions,
$\mu_{\phi_1}$ and $\mu_{\phi_2}$, extracted in this way.
For the surface density, we subtract an estimate of the background surface
density by \todo{AB}.
To estimate the stream width, $w$, we compute the median absolute deviation,
$\textrm{MAD}$, and use this to estimate a robust standard deviation, $w = 1.5
\times \textrm{MAD}$.

\todo{Comment on density variations, stream width variations. Do we want a
$\Delta \phi_2$ plot (relative to mean stream track) with 15th/85th
percentiles?}

- first measure density variations in background subtracted profile
- find maximum -- morphology as expected from a progenitor in the final stages of dissolution (cit)
- several prominent gaps: density contrast, width
- some (all?) seen in earlier data
- but not related to dust

@@

To initialize the model stream generation, we assume that the present day (i.e.
final) position of the progenitor system is at $\phi_1 = XX^\circ$.
We integrate backwards for $4~\textrm{Gyr}$ using the orbital parameters at this
location using the Leapfrog integration scheme with a timestep of
$0.5~\textrm{Myr}$, then begin the model stream generation using the Lagrange
point stripping method discussed above.
Stream particles are release at each timestep of the progenitor orbit
integration, resulting in 16,000 stream particles by the end of the model stream
generation.
We assume an initial progenitor mass of $M=10^5~\msun$, and linearly decrease
the mass of the progenitor until present day, where we assume $M = 0~\msun$
(i.e. fully disrupted).
Stream properties computed from this resulting model stream are plotted in
\figurename~\ref{fig:track-and-model} as \todo{XX}.

@@

\appendix
\section{Completeness and the \gaia\ scanning pattern}
\label{sec:completeness}

>>>>

% Make the point that if we can measure distance to spur, stream, RVs - measure energy offset which doesn't depend on time, like gap size

% Make sure to include simple equations so can rescale

% Note that epicyclic density variations naturally in, but caveat uniform release times

@@

To study the global properties of the stream as a function of stream longitude,
we extract stream stars around a low-order polynomial model for the median
stream latitude, $\phi_2$, as a function of stream longitude, $\phi_1$.
In detail, we find stream ridge points by computing a running median in $\phi_2$
in steps of $2^\circ$ with a window size of $4^\circ$ along $\phi_1$, then fit a
second-order polynomial to the ridge points to find $\phi_{2,
\textrm{track}}(\phi_1)$.
We define the stream as the region within $\left| \phi_2 - \phi_{2,
\textrm{track}}(\phi_1) \right| < 0.75^\circ$.
\figurename~\ref{fig:track-and-model} (top panel) again shows the
high-confidence stream stars, and the two curved lines (blue) show the adopted
upper and lower boundary of the stream region.

We extract properties of the stream from the region defined above as a function
of $\phi_1$ by computing stream properties in overlapping windows:
we set the window size to $4^\circ$ in $\phi_1$, and shift the window center by
$1^\circ$ in longitude for each measurement.
\figurename~\ref{fig:track-and-model} shows the median latitude, $\phi_2$,
stream surface density, $XX$, stream width, $XX$, and median proper motions,
$\mu_{\phi_1}$ and $\mu_{\phi_2}$, extracted in this way, and empirical scatters
for each estimated from each window (blue lines and shaded regions).
For the surface density estimate, we subtract an estimate of the background
surface density computed from the area outside of the bounded stream region.

Clear density variations, also apparent in the 2D positions of the sources,
manifest as sharp features in the surface density estimates along the stream;
these are not a result of variable dust extinction or completeness variations
from the \gaia\ scanning pattern.
First, the visible portions of the GD-1 stream are located at high Galactic
latitudes ($b > 20^\circ$), and we therefore do not expect significant dust
extinction or variations in extinction within the stream region.
For stream longitudes $-60^\circ < \phi_1 < 10^\circ$ and $-1^\circ < \phi_2 <
1^\circ$, the median $V$-band extinction is $\textrm{med}\left(A_V\right) =
0.04~\textrm{mag}$, and the dispersion in $A_V$ in this region is $\sigma_{A_V}
\approx 0.03~\textrm{mag}$, as computed from the Schlegel-Finkbeiner-Davis
extinction map (\citealt{Schlegel:1998}).
For longitudes $\phi_1 < -60^\circ$, as the stream approaches the Galactic disk,
dust extinction becomes more appreciable, and the stream stars are harder to
select apart from the background.
Second, \todo{what to say about the Gaia scanning law...}

The density variations likely represent real morphological changes along the
GD-1 stream, as have been suggested before from binned samples of
color-magnitude-selected GD-1 stars (\citealt{Carlberg:2013, DeBoer:2018}).
The deep under-densities we see from the kinematically-cleaned sample correspond
to previously reported gaps in the stream.
There is also a well-defined, sharp over-density in the surface density profile
close to $\phi_1 \approx -13^\circ$ with roughly symmetric under-densities on
either side:
this is suggestive of a progenitor system in the final stages of dissolution
(\todo{cite who}).
\todo{Table of locations, sizes, and density constrasts for the features?}

@@

To initialize the model stream generation, we assume that the present day (i.e.
final) position of the progenitor system is at $\phi_1 = -13.5^\circ$, the peak
of the observed surface density profile.
We integrate backwards for $4~\textrm{Gyr}$ using the orbital parameters at this
location using the Leapfrog integration scheme with a timestep of
$0.5~\textrm{Myr}$, then begin the model stream generation using the Lagrange
point stripping method discussed above.
Stream particles are release at each timestep of the progenitor orbit
integration, resulting in 16,000 stream particles by the end of the model stream
generation.
We assume an initial progenitor mass of $M=10^5~\msun$, and linearly decrease
the mass of the progenitor until present day, where we assume $M = 0~\msun$
(i.e. fully disrupted).
Stream properties computed from this resulting model stream are plotted in
\figurename~\ref{fig:track-and-model} as \todo{XX}.

@@

% \appendix
% \section{Completeness and the \gaia\ scanning pattern}
% \label{sec:completeness}

% \figurename~\ref{fig:XX} (XX panel) shows the $V$-band extinction
% in the region around the GD-1 stream, computed from the
% Schlegel-Finkbeiner-Davis extinction map (\cite{Schlegel:1998}; hereafter SFD).
